# Practical Machine Learning Course Project

The goal of this project is to examine and predict the effects of contributing variables to the "classe" variable in the training set.

## Uploading Required Packages

I begin this project with uploading the following packages into R:

```{r}
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)
```

## Importing Data

After downloading data sets from the links provided by Coursera, they have to be read by R:

```{r}
test <- read.csv("pml-testing.csv", header=T, na.strings = c("NA", "#DIV/0!", ""))
train <- read.csv("pml-training.csv", header=T, na.strings = c("NA", "#DIV/0!", ""))
```

The next step is to look closely at the obtained data:

```{r, results='hide'}
str(train)
head(train)
str(test)
head(test)
```

## Cleaning Data

To exclude all of variables with at least one "NA" from the data sets, the following codes should yield the desired results:

```{r}
train2 <- train[, apply(train, 2, function(x) !any(is.na(x)))] 
test2 <- test[, apply(test, 2, function(x) !any(is.na(x)))] 
```

In addition, columns one to eight can be removed from the data sets: 

```{r}
train3 <- train2 [,-c(1:8)]
test3 <- test2 [,-c(1:8)]
```

## Spliting Data

The next step is to split train data set into two new data sets: training and testing. I used 75% for training data set size and 25% for testing: 

```{r}
set.seed(65478)
inTrain <- createDataPartition(y=train3$classe, p=0.75, list=FALSE)
training  <- train3[inTrain,]
testing  <- train3[-inTrain,]
```

## Modeling

Alright! The data is processed and ready to get analyzed. I will start the data analysis process with fitting a CART model:

```{r}
set.seed(654789)
cartmodel <- rpart(classe ~ ., data=training, method="class")
prp(cartmodel)
```

Then, developing a random forest model may be helpful:

```{r}
set.seed(6547891)
rfmodel <- randomForest(classe ~. , data=training)
```

## Results

To select the best model for this data, it is necessary to run a sensitivity analysis. Taking alook at the prediction rates extracted from both CART and random forests should be useful:

```{r}
precart <- predict(cartmodel, testing, type = "class")
confusionMatrix(precart, testing$classe)
prerf <- predict(rfmodel, testing, type = "class")
confusionMatrix(prerf, testing$classe)
```

## Conclusion

Based on the results, the performance of random forests model is great with a 99.53% accuracy. The accuracy for the CART model is 69.88% that is acceptable too. Thus, I will choose the RF model.

## Out-of-Sample Error

Based on the results from random forest model, the out of sample error is 0.0047 (1-0.9953=0.0047) or 0.47%. Pretty Cool!
